A Polls-of-Polls Forecast for the 2024 United States Presidential Election:Analysis of Support for Harris Factoring Pollster Differences

### 2024-11-03 paper/paper.qmd
Introduction + Data (overivew and methodology) completed

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Jiwon Choi
  - Kevin Roe
thanks: "Code and data are available at: [https://github.com/jwonc4602/2024_US_Election_Forecast](https://github.com/jwonc4602/2024_US_Election_Forecast)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

@@

We have used a poll-of-polls of the upcoming US presidential election dataset obtained from FiveThirtyEight [@fivethirtyeight_polls]. Data was collected and analyzed using R statistical programming software [@citeR], with additional packages like tidyverse [@tidyverse], rstanarm [@rstanarm] knitr [@citeKnitr], here [@here], and many others for support.

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

@@

## Predictor variables

@@

# Model
For our analysis, we employ a Bayesian generalized linear model (GLM) to forecast the popular vote percentage for Kamala Harris. This approach allows us to capture polling characteristics and account for known variations between pollsters, population types, and the recency of polling data. By incorporating these factors, we aim to model a nuanced estimate of Harris's projected popular vote.

@@

# Results
To assess model reliability, we examined several key diagnostics. Convergence metrics, such as Rhat values, were close to 1 for all parameters, indicating strong convergence. Additionally, the effective sample size (n_eff) was high across parameters, suggesting low autocorrelation and contributing to model stability. For validation, we conducted out-of-sample testing and calculated the Root Mean Square Error (RMSE) to assess predictive accuracy. This test data, which includes polls not used in training, provides an unbiased estimate of model performance.

@@

Table @tbl-model-coefficients presents the estimated coefficients for the predictors in our GLM model. These coefficients fit into the GLM equation, allowing us to interpret the impact of each predictor on Harris’s predicted vote percentage. Positive values indicate a higher predicted percentage, while negative values indicate a decrease. Key predictors, such as specific pollsters and population types, show distinct effects on the forecasted outcomes. This table is produced using kable from knitr for enhanced readability.

@@

Figure @fig-coefficient-estimates visually represents the model coefficients, with error bars indicating the confidence interval for each estimate. Positive coefficients suggest that specific pollsters or demographic characteristics increase Harris’s predicted vote share. For example, the 'CES / YouGov' pollster shows a significant positive impact, while 'Siena/NYT' has a slight negative effect. These error bars help contextualize the reliability of each predictor.

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point {#sec-second-point}

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point {#sec-third-point}

@@

## Weaknesses and Next Steps {#sec-weaknesses}

>>>>

---
title: "A Polls-of-Polls Forecast for the 2024 United States Presidential Election:Analysis of Support for Harris Factoring Pollster Differences"
author: 
  - Jiwon Choi
  - Kevin Roe
thanks: "Code and data are available at: [https://github.com/jwonc4602/2024_US_Election_Forecast](https://github.com/jwonc4602/2024_US_Election_Forecast)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

The United States Presidential Election is one of the most consequential events of 2024. As a key country in international relations and the largest economy in the world, the results of the U.S election determines the country's domestic and foreign policy for the next four years, which will have a significant impact on important initiatives such as tackling climate change or international conflicts. In anticipation of the 2024 U.S election, this paper aims to predict the possible outcomes of the election by analyzing the level of support that Kamala Harris will gain.

We forecast support for Kamala Harris based on polling results and apply a Bayesian generalized linear model. The main parameter of interest is the proportion of vote or support that Harris received in surveys, which is traced over time. By considering the results from different poll-making organizations and other demographic factors, our objective is to correct for variation across different voter bases with

Factoring in the different results of poll-making organizations, our model found that specific pollsters or demographic characteristics increase or decrease Harris' predicted voter share. Understanding the trajectory of the US might make in foreign policy, economic or global politics prediction allows global stakeholders to take preemptive measures for policy changes of the newly elected government. Thus, this study is a strong tool to navigate uncertainty around the 2024 election.

The paper is structured as follows: @sec-data and @sec-model explores the data and methodology used, highlighting the filtering and modeling techniques applied to the data; @sec-results presents the results from the Bayesian generalized linear model; @sec-discussion discusses the broader implications of our findings; and @sec-appendix highlights YouGov's methodology, and outlines an idealized survey methodology *SHOULD WE ADD MODEL DIAGNOSTICS HERE?*

@@

We have used a poll-of-polls of the upcoming US presidential election dataset obtained from FiveThirtyEight [@fivethirtyeight_polls]. Data was collected and analyzed using R statistical programming software [@citeR], with additional packages like tidyverse [@tidyverse], rstanarm [@rstanarm] knitr [@citeKnitr], here [@here], and many others for support. The dataset includes a wide range of poll results from various polls, with keyv ariables such as pollster, sample size, percentage of support for Harris, and the date conducted for polls.

To ensure high data quality, we filtered the dataset to include polls with a numeric grade of 2.7 or above and focused only on polls conducted after July 21, 2024, the date of Harris' declaration of candidacy. The filter helped limit bias from older polls from when Joe Biden was the Democratic nominee, which may not reflect the current voter sentiment.

In performing the analysis, we used several R pacakages. We used tidyverse [@tidyverse] for data manipulation and visualization, and rstanarm [@rstanarm] for Bayesian modeling. We also used modelsummary [@modelsummary] to generate model summaries.To visualize results, we used ggplot2 [@ggplot2] and kable [@kable] *ADD KABLE CITATION?* to format tables for presentation.

## Measurement {#sec-data-measurement}

FiveThirtyEight aggregates various poll results from national and state-wide polls, showing 16867 observations [@fivethirtyeight_polls]. The poll takes a sample of the electorate and asks for the voters' candidate of choice. By factoring polls from a state and national level, FiveThirtyEight aims to represnet public perception on the two candidates during the election to predict the election's outcome.

Each poll aims to predict an actual event. The raw data is susceptible because each pollster has varying polling methods. There are other limitations such as sampling error, response error, or distortion from participants misunderstanding the question.

Beyond filtering for high-quality pollsters, we filtered for various criteria to prepare the analysis data.From the raw data, we filtered out any missing numbers, cleaned names, and removed all polls on non-Democrat or Republican candidates, effectively on comparing Harris and Trump. After determining state-level analysis for electoral votes and close races, we created different datasets for national and state level polls. Regardless, selection bias and sampling bias is still a concern because polls only represent a part of the population. Finally, polling methodology difference can introduce bias in the study. Overall, we use the data from individual responses to election polls, and filter the data for analysis to make a prediction on who will win the popular vote based on data from the pollsters.

## Outcome variables {#sec-data-outcome}

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
- include summary statistics 
- for every variable in your dataset that is of interest, there needs to be graphs, explanations and tables

@@

## Predictor variables {#sec-data-predictor}

@@

# Model {#sec-model}
For our analysis, we employ a Bayesian generalized linear model (GLM) to forecast the popular vote percentage for Kamala Harris. This approach allows us to capture polling characteristics and account for known variations between pollsters, population types, and the recency of polling data. By incorporating these factors, we aim to model a nuanced estimate of Harris's projected popular vote.

@@

# Results {#sec-results}
To assess model reliability, we examined several key diagnostics. Convergence metrics, such as Rhat values, were close to 1 for all parameters, indicating strong convergence. Additionally, the effective sample size (n_eff) was high across parameters, suggesting low autocorrelation and contributing to model stability. For validation, we conducted out-of-sample testing and calculated the Root Mean Square Error (RMSE) to assess predictive accuracy. This test data, which includes polls not used in training, provides an unbiased estimate of model performance.

@@

@tbl-model-coefficients presents the estimated coefficients for the predictors in our GLM model. These coefficients fit into the GLM equation, allowing us to interpret the impact of each predictor on Harris’s predicted vote percentage. Positive values indicate a higher predicted percentage, while negative values indicate a decrease. Key predictors, such as specific pollsters and population types, show distinct effects on the forecasted outcomes. This table is produced using kable from knitr for enhanced readability.

@@

@fig-coefficient-estimates visually represents the model coefficients, with error bars indicating the confidence interval for each estimate. Positive coefficients suggest that specific pollsters or demographic characteristics increase Harris’s predicted vote share. For example, the 'CES / YouGov' pollster shows a significant positive impact, while 'Siena/NYT' has a slight negative effect. These error bars help contextualize the reliability of each predictor.

# Discussion {#sec-discussion}

## How has the Forecast Changed Over Time? {#sec-first-point}

asdfasdfasdfasdf

## Who is Favoured to Win Each State? {#sec-second-point}

## What are the Closest Races Across States {#sec-third-point}

@@

## Weaknesses and Next Steps {#sec-weaknesses}
- if filter thorugh state, data not collected easily
